Bioverse: Potentially Observable Exoplanet Biosignature Patterns Under the UV Threshold Hypothesis for the Origin of Life

### 2025-03-08 src/tex/ms.tex
fix a few formatting issues

\rev{\citet{Ranjan2017b} speculated that UV photochemistry might be rate-limiting for abiogenesis, particularly on planets orbiting M dwarfs, due to their lower baseline UV fluxes.
This could delay abiogenesis by orders of magnitude, resulting in a continuous dependence of abiogenesis likelihood on UV flux.
However, recent studies challenge this view for the cyanosulfidic scenario.
\citet{Rimmer2021}  calculated photochemical timescales on early Earth at 180-300 hours (7.5-12.5 days), significantly shorter than the timescale for stochastic geological events also required for the scenario~\citep{Rimmer2023}.
Even with 1000x slower photochemistry on M-dwarf planets, prebiotic photochemistry like sulfite photolysis would still occur on a geologically negligible timescale of 20-30 years, meaning that the photochemistry is unlikely to be rate limiting compared to stochastic geological processes, justifying the step-function model.
}

@@

\rev{
To assess the implications of a more conservative abiogenesis timescale, we consider a scenario in which a planet's minimum time required to be both in the \gls{HZ} and having above-threshold \gls{NUV} fluxes before life can emerge is increased to \(\Delta T_{\text{min}} = 100\) Myr, rather than 1 Myr as assumed above.
Figures~\ref{fig:inhabited_dT100}, \ref{fig:surveys_dT100}, and \ref{fig:powergrid_dT100} show the results of our \bioverse\ simulations under this assumption.
}

\rev{
The overall impact differs significantly between M dwarf and FGK dwarf planets.
 While M dwarf planets remain largely unaffected, the fraction of inhabited FGK dwarf planets decreases sharply due to the limited overlap between the time of high UV flux and habitable conditions (compare Figure~\ref{fig:hz_nuv_evo}).
Even under the optimistic assumption that the probability of abiogenesis is unity (\( f_{\text{life}} = 1 \)), the fraction of inhabited planets under the UV Threshold Hypothesis declines rapidly with increasing UV flux thresholds (Figure~\ref{fig:inhabited_dT100}).
}
\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{figures/inhabited_aafo_nuv_thresh_deltatmin100.pdf}
    \caption{\rev{Fraction of inhabited planets as a function of NUV threshold flux for \(\Delta T_{\text{min}} = 100\) Myr. The decrease in inhabited planets is more pronounced for FGK stars, narrowing the viable range of UV Thresholds compared to shorter origins timescales.}}
    \label{fig:inhabited_dT100}
\end{figure}
\begin{figure*}
    \centering
    \includegraphics[width=0.9\linewidth]{figures/surveys_FGKM_dT100.pdf}
    \caption{\rev{Simulated survey results under the assumption of \(\Delta T_{\text{min}} = 100\) Myr. The fraction of inhabited FGK planets is reduced to zero, while the fraction of inhabited M dwarf planets remains largely unaffected.}}
    \label{fig:surveys_dT100}
\end{figure*}

\rev{
The impact of a longer origins timescale on testing the UV Threshold Hypothesis becomes evident when we repeat our example survey ($f_{\text{life}} = 0.8$, $F_{\text{NUV},\text{min}} = \SI{300}{\erg\per\second\per\centi\meter\squared}$) under this assumption (Figure~\ref{fig:surveys_dT100}):
While the survey of M dwarf planets remains largely unaffected, the fraction of inhabited FGK planets is reduced to zero, making a conclusive test of the UV Threshold Hypothesis impossible.
}
\begin{figure*}
    \centering
    \includegraphics[width=0.9\linewidth]{figures/powergrid_dT100.pdf}
    \caption{\rev{Statistical power for different abiogenesis rates and NUV threshold fluxes under the assumption of \(\Delta T_{\text{min}} = 100\) Myr. The longer origins timescale makes a test of the UV Threshold Hypothesis impossible for FGK planets, while M dwarf planets remain viable targets.}}
    \label{fig:powergrid_dT100}
\end{figure*}
\rev{
Figure \ref{fig:powergrid_dT100} demonstrates that this result is independent of the threshold \gls{NUV} flux and the intrinsic abiogenesis rate.
Given the lack of constraints on the timescale of abiogenesis, these findings highlight the advantages of focusing on M dwarf planets when searching for biosignatures.
If life emerges slowly, FGK planets may rarely, if ever, reach the inhabited stage, whereas M dwarf planets remain viable targets for future surveys.
If biosignatures were nevertheless detected on FGK planets despite the constraints imposed by a longer abiogenesis timescale, it would suggest that either the UV Threshold Hypothesis, as formulated here, is incorrect or that the timescale of abiogenesis is shorter than $\mathcal{O}(\SI{100}{\mega\year})$.
}
\end{document}

>>>>

\rev{\citet{Ranjan2017b} speculated that UV photochemistry might be rate-limiting for abiogenesis, particularly on planets orbiting M dwarfs, due to their lower baseline UV fluxes.
This could delay abiogenesis by orders of magnitude, resulting in a continuous dependence of abiogenesis likelihood on UV flux.
However, recent studies challenge this view for the cyanosulfidic scenario.
\citet{Rimmer2021} calculated photochemical timescales on early Earth at 180--300 hours (7.5--12.5 days), significantly shorter than the timescale for stochastic geological events also required for the scenario~\citep{Rimmer2023}.
Even with 1000x slower photochemistry on M-dwarf planets, prebiotic photochemistry like sulfite photolysis would still occur on a geologically negligible timescale of 20--30 years, meaning that the photochemistry is unlikely to be rate limiting compared to stochastic geological processes, justifying the step-function model.
}

@@

\rev{
To assess the implications of a more conservative abiogenesis timescale, we consider a scenario in which a planet's minimum time required to be both in the \gls{HZ} and having above-threshold \gls{NUV} fluxes before life can emerge is increased to $\Delta T_{\mathrm{min}} = 100$ Myr, rather than 1 Myr as assumed above.
Figures~\ref{fig:inhabited_dT100}, \ref{fig:surveys_dT100}, and \ref{fig:powergrid_dT100} show the results of our \bioverse\ simulations under this assumption.
}

\rev{
The overall impact differs significantly between M dwarf and FGK dwarf planets.
 While M dwarf planets remain largely unaffected, the fraction of inhabited FGK dwarf planets decreases sharply due to the limited overlap between the time of high UV flux and habitable conditions (compare Figure~\ref{fig:hz_nuv_evo}).
Even under the optimistic assumption that the probability of abiogenesis is unity ($ f_{\mathrm{life}} = 1 $), the fraction of inhabited planets under the UV Threshold Hypothesis declines rapidly with increasing UV flux thresholds (Figure~\ref{fig:inhabited_dT100}).
}
\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{figures/inhabited_aafo_nuv_thresh_deltatmin100.pdf}
    \caption{\rev{Fraction of inhabited planets as a function of NUV threshold flux for $\Delta T_{\mathrm{min}} = 100$ Myr. The decrease in inhabited planets is more pronounced for FGK stars, narrowing the viable range of UV Thresholds compared to shorter origins timescales.}}
    \label{fig:inhabited_dT100}
\end{figure}
\begin{figure*}
    \centering
    \includegraphics[width=0.9\linewidth]{figures/surveys_FGKM_dT100.pdf}
    \caption{\rev{Simulated survey results under the assumption of $\Delta T_{\mathrm{min}} = 100$ Myr. The fraction of inhabited FGK planets is reduced to zero, while the fraction of inhabited M dwarf planets remains largely unaffected.}}
    \label{fig:surveys_dT100}
\end{figure*}

\rev{
The impact of a longer origins timescale on testing the UV Threshold Hypothesis becomes evident when we repeat our example survey ($f_{\mathrm{life}} = 0.8$, $F_{\mathrm{NUV},\mathrm{min}} = \SI{300}{\erg\per\second\per\centi\meter\squared}$) under this assumption (Figure~\ref{fig:surveys_dT100}):
While the survey of M dwarf planets remains largely unaffected, the fraction of inhabited FGK planets is reduced to zero, making a conclusive test of the UV Threshold Hypothesis impossible.
}
\begin{figure*}
    \centering
    \includegraphics[width=0.9\linewidth]{figures/powergrid_dT100.pdf}
    \caption{\rev{Statistical power for different abiogenesis rates and NUV threshold fluxes under the assumption of $\Delta T_{\mathrm{min}} = 100$ Myr. The longer origins timescale makes a test of the UV Threshold Hypothesis impossible for FGK planets, while M dwarf planets remain viable targets.}}
    \label{fig:powergrid_dT100}
\end{figure*}
\rev{
Figure \ref{fig:powergrid_dT100} demonstrates that this result is independent of the threshold \gls{NUV} flux and the intrinsic abiogenesis rate.
Given the lack of constraints on the timescale of abiogenesis, these findings highlight the advantages of focusing on M dwarf planets when searching for biosignatures.
If life emerges slowly, FGK planets may rarely, if ever, reach the inhabited stage, whereas M dwarf planets remain viable targets for future surveys.
If biosignatures were nevertheless detected on FGK planets despite the constraints imposed by a longer abiogenesis timescale, it would suggest that either the UV Threshold Hypothesis, as formulated here, is incorrect or that the timescale of abiogenesis is shorter than $\mathcal{O}(\SI{100}{\mega\year})$.
}
\end{document}
